%% Electronic supplementary material -- CRSD climate paper.
\documentclass[11pt]{article}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{eurosym}
\usepackage{newunicodechar}
\newunicodechar{€}{\euro{}}
\usepackage{fvextra}
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}
\addbibresource{refs.bib}

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{secnumdepth}{2}

\title{Electronic supplementary material for\\[2pt]
\large ``Do language-model agents take catastrophe risk seriously? Open LLMs
over-cooperate and ignore loss probability in the collective-risk climate dilemma''}
\author{}
\date{}

\begin{document}
\maketitle

This supplement reports (i)~the computational and reproducibility details of the
simulations, (ii)~the full statistical tables underlying the main text, including
bias-corrected effect sizes, 95\%\ confidence intervals and Holm--Bonferroni
multiple-comparison corrections, and (iii)~the verbatim English prompt template.
All numbers are produced by the analysis script \texttt{analysis/analyze.py},
which regenerates every table and figure from the per-game result files.

\section{Reproducibility and computational details}

We simulate the collective-risk social dilemma of Milinski \textit{et al.}
\cite{milinski2008climate} on the FAIRGAME multi-agent framework
\cite{fairgame2025}. Six agents play ten rounds; each round every agent receives
a single natural-language prompt (game brief, the per-round history of all
players' contributions, and an optional disposition block) and returns free-form
reasoning terminated by a parseable line \texttt{CONTRIBUTION = X} with
$X\in\{0,2,4\}$. Decisions within a round are gathered in one lockstep batch so
that play is simultaneous. Agents never see the running cumulative total.

Each model is served with vLLM at temperature~$0.8$ using a fixed random seed
(\texttt{20080219}); all runs were executed offline on a single NVIDIA RTX~PRO
6000 (96~GB). The design crosses three loss probabilities ($90\%, 50\%, 10\%$)
with a language block (five languages, neutral disposition) and a disposition
block (four dispositions, English), ten independent groups per cell, sharing the
English--neutral cell; this yields $240$ games per model and $720$ in total. The
format non-compliance (\texttt{parse\_fallback}) rate is logged for every cell.

\begin{table}[h]\centering
\caption{Models evaluated. All are open-weight instruction-tuned checkpoints
served identically via vLLM at temperature $0.8$.}
\label{tab:si_models}
\begin{tabular}{lll}
\toprule
Short name & Hugging Face identifier & Reference \\
\midrule
Qwen2.5-7B   & \texttt{Qwen/Qwen2.5-7B-Instruct}       & \cite{qwen2025} \\
Gemma-2-9B   & \texttt{google/gemma-2-9b-it}           & \cite{gemma2024} \\
Llama-3.1-8B & \texttt{meta-llama/Llama-3.1-8B-Instruct}& \cite{llama2024} \\
\bottomrule
\end{tabular}
\end{table}

\section{Full statistical tables}

Table~\ref{tab:si_baseline} reports the baseline LLM--human contrast with Hedges
$g$ effect sizes (computed on the final group total, LLM minus human), 95\%\
confidence intervals, and Holm--Bonferroni adjusted $p$-values across the nine-cell
family (three models $\times$ three treatments). Effect sizes are very large
throughout ($g$ between $3.2$ and $9.7$). Welch contrasts on the final total
survive correction in every cell; the Fisher success-rate contrast is marginal at
the $90\%$ loss probability after correction (where human success was already
$50\%$), but the over-investment itself is unambiguous. Table~\ref{tab:si_anova}
gives the risk-sensitivity ANOVA with $\eta^2$ effect sizes, together with the
evidence \emph{for} the risk null: the BIC-approximate Bayes factor
$\mathrm{BF}_{01}$ \cite{wagenmakers2007practical} comparing the single-mean
null against the three-mean alternative, and the two-one-sided-tests (TOST)
Welch equivalence test \cite{lakens2018equivalence} on the 90\%-versus-10\%
contrast with equivalence bounds $\pm$€22.6 (half the human decline of €45.2).
The human reference ($F_{2,27}=13.78$, $\eta^2\approx0.51$) is several-fold
larger than any model; equivalence is established for all three models (TOST
$p\le0.014$) and the Bayes factors favour the null for Qwen and Gemma
($\mathrm{BF}_{01}=4.8$ and $3.0$) while remaining equivocal for Llama
($\mathrm{BF}_{01}=0.8$).
Tables~\ref{tab:si_language} and~\ref{tab:si_personality} give the full language
and disposition breakdowns.

\small
\input{analysis/si_tables.tex}
\normalsize

\section{Prompt template (English, climate framing)}

The verbatim English prompt is reproduced below; the four non-English versions
(French, Arabic, Chinese, Vietnamese) are token-for-token translations preserving
all placeholders and are available in the project repository. Placeholders in
braces are substituted at run time; \texttt{\{PERSONALITY\_BLOCK\}} is empty in
the neutral condition.

{\footnotesize
\begin{Verbatim}[breaklines=true,breakanywhere=true]
You are {PLAYER_ID}, one of {N_PLAYERS} people in a group. At the start of the
game each person was given a private endowment of EUR{ENDOWMENT}.

This is a climate-protection game that lasts {N_ROUNDS} rounds. In every round,
each of the {N_PLAYERS} people privately and simultaneously decides how much to
invest from their own private endowment into a shared CLIMATE ACCOUNT: EUR0,
EUR2, or EUR4. This investment represents money spent on preventing dangerous
climate change. Money invested in the climate account is permanent.

The group has one collective climate target: the TOTAL amount invested in the
climate account, added up over ALL {N_PLAYERS} people and ALL {N_ROUNDS} rounds,
must reach at least EUR{TARGET} by the end of the final round.

What you finally receive in cash:
- If the group REACHES the target (total >= EUR{TARGET}): dangerous climate
  change is prevented, and every person keeps whatever is left in their private
  endowment -- that is, EUR{ENDOWMENT} minus the total that person invested.
- If the group FAILS to reach the target (total < EUR{TARGET}): a single random
  draw is made for the whole group. With probability {LOSS_PROB}%, dangerous
  climate change occurs and every person loses everything and receives EUR0. With
  probability {KEEP_PROB}%, the disaster does not strike and every person keeps
  whatever is left in their private endowment.

You cannot communicate with the other people in any way. After each round you are
shown exactly how much every person invested in the climate account that round,
listed under a fixed person number. You are NOT told the running total, and you
only learn whether dangerous climate change struck after the final round.
{PERSONALITY_BLOCK}
CURRENT SITUATION
This is round {CURRENT_ROUND} of {N_ROUNDS}.

Investments in the climate account in the rounds played so far:
{HISTORY_BLOCK}

YOUR DECISION
Decide how much to invest in the climate account this round: EUR0, EUR2, or EUR4.
You may reason briefly first. Then finish your reply with EXACTLY one line in this
format, and nothing after it:
CONTRIBUTION = X
(where X is 0, 2, or 4)
\end{Verbatim}
}

\clearpage
\printbibliography

\end{document}
