%% Electronic supplementary material -- PGG punishment paper.
\documentclass[11pt]{article}
\usepackage[a4paper,margin=2cm]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{fvextra}
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}
\addbibresource{refs.bib}

\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{secnumdepth}{2}

\title{Electronic supplementary material for\\[2pt]
\large ``Antisocial but inconsequential: large language model agents punish
cooperators yet neither sustain cooperation nor reproduce its cross-cultural
geography''}
\author{}
\date{}

\begin{document}
\maketitle

This supplement reports (i)~the computational and reproducibility details, (ii)~the
full statistical tables underlying the main text, including effect sizes,
confidence intervals, the complete deviation-binned punishment, the per-society
breakdown and the dispositional analysis, and (iii)~the verbatim English prompt
templates for both stages. All numbers are produced by \texttt{analysis/analyze.py}.

\section{Reproducibility and computational details}

We simulate the public-goods game with costly peer punishment of Herrmann
\textit{et al.}\ \cite{herrmann2008antisocial} on the FAIRGAME framework
\cite{fairgame2025}. Four agents are matched as partners for ten periods. Each
period every agent receives a 20-token endowment and a contribution prompt, and
returns reasoning terminated by \texttt{CONTRIBUTION = X} ($0\le X\le20$); in
treatment~P a second prompt elicits a punishment vector \texttt{DEDUCT: A=a B=b
C=c} ($0\le a,b,c\le10$). Co-players are re-labelled every period to block
identity tracking, and decisions within a stage are gathered in one lockstep
batch.

Each model is served with vLLM at temperature~$0.8$ using a fixed random seed
(\texttt{20080307}); all runs were executed offline on a single NVIDIA RTX~PRO
6000 (96~GB). The design crosses two treatments (N, P) with a language block
(five languages, neutral disposition), a disposition block (four dispositions,
English) and a society block (sixteen neutral city personas, English), ten
independent groups per cell, sharing the English--neutral--no-society cell; this
yields $480$ games per model and $1440$ in total. Punishment is classified by the
contribution deviation between target and punisher: deviation $\geq 0$ (the target
gave at least as much as the punisher) is antisocial, deviation $<0$ is punishment
of free-riding. The format non-compliance rate is logged for every cell.

\begin{table}[h]\centering
\caption{Models evaluated. All are open-weight instruction-tuned checkpoints
served identically via vLLM at temperature $0.8$.}
\label{tab:si_models}
\begin{tabular}{lll}
\toprule
Short name & Hugging Face identifier & Reference \\
\midrule
Qwen2.5-7B   & \texttt{Qwen/Qwen2.5-7B-Instruct}       & \cite{qwen2025} \\
Gemma-2-9B   & \texttt{google/gemma-2-9b-it}           & \cite{gemma2024} \\
Llama-3.1-8B & \texttt{meta-llama/Llama-3.1-8B-Instruct}& \cite{llama2024} \\
\bottomrule
\end{tabular}
\end{table}

\section{Full statistical tables}

Table~\ref{tab:si_punish} reports the punishment effect on cooperation (P$-$N) with
Hedges $g$ and rank-biserial $r$ effect sizes, Holm-adjusted Mann--Whitney
$p$-values and 95\%\ confidence intervals: only Gemma differs from zero, and in the
negative direction ($g=-2.93$, $r=-0.94$). Table~\ref{tab:si_deviation} gives the
full deviation-binned punishment for all three models, showing that only Gemma
reproduces the human monotone free-riding gradient. Table~\ref{tab:si_cross}
reports the cross-societal correlations with Fisher-$z$ confidence intervals; none
approaches the human $\rho\approx-0.90$. Table~\ref{tab:si_society} gives the full
per-society contributions for the primary model against the human pools, and
Table~\ref{tab:si_personality} the dispositional breakdown, in which a
\emph{vengeful} prompt raises the antisocial share without restoring the
institutional function of punishment.

\small
\input{analysis/si_tables.tex}
\normalsize

\section{Prompt templates (English)}

The verbatim English contribution-stage and punishment-stage prompts are
reproduced below; the four non-English versions are token-for-token translations
preserving all placeholders, available in the project repository. Placeholders in
braces are substituted at run time; \texttt{\{SOCIETY\_BLOCK\}} and
\texttt{\{PERSONALITY\_BLOCK\}} are empty in the neutral, no-society condition.

\subsection*{Contribution stage}
{\footnotesize
\begin{Verbatim}[breaklines=true,breakanywhere=true]
You are {PLAYER_ID}, one of {N_PLAYERS} members of a group. The same {N_PLAYERS}
members stay together for the whole game.

The game lasts {N_ROUNDS} periods. At the start of EVERY period each member
receives a fresh endowment of {ENDOWMENT} tokens. In each period you privately and
simultaneously decide how many of your {ENDOWMENT} tokens to put into a shared
group project; you keep the rest in your private account.

How tokens become income each period:
- Every token you keep in your private account is worth 1 token of income to you.
- Every token placed in the group project is shared among all {N_PLAYERS} members:
  each member receives {MPCR} tokens of income for EACH token in the project, no
  matter who put it there.

You contribute {ACTION_DESC} to the group project this period.

The other members are shown to you only under temporary labels ({OTHER_LABELS}).
These labels are reassigned every period and do not tell you who is who across
periods. You cannot communicate in any way, and all decisions are anonymous.
{TREATMENT_BLOCK}
After each period you are shown how many tokens each member placed in the project
that period, the total in the group project, and your own income for that period.
{SOCIETY_BLOCK}
{PERSONALITY_BLOCK}
CURRENT SITUATION
This is period {CURRENT_ROUND} of {N_ROUNDS}.

Results of the periods played so far:
{HISTORY_BLOCK}

YOUR DECISION
Decide how many of your {ENDOWMENT} tokens to place in the group project this
period ({ACTION_DESC}). You may reason briefly first. Then finish your reply with
EXACTLY one line in this format, and nothing after it:
CONTRIBUTION = X
(where X is an integer from {CONTRIB_MIN} to {CONTRIB_MAX})
\end{Verbatim}
}

\subsection*{Punishment stage (treatment P)}
{\footnotesize
\begin{Verbatim}[breaklines=true,breakanywhere=true]
You are {PLAYER_ID}. This is the second stage of period {CURRENT_ROUND} of
{N_ROUNDS}.

This period you placed {MY_CONTRIB} tokens in the group project. The total placed
by all members was {GROUP_PROJECT_TOTAL}, and your income from the first stage
this period is {MY_INCOME_THIS_PERIOD} tokens.

Here is how much each of the other members placed in the project this period:
{OTHERS_BLOCK}

You may now assign deduction points to each of these members, simultaneously and
anonymously. For EACH member you choose a whole number of deduction points from 0
to {MAX_PUNISH}.
- Each deduction point you assign costs YOU {PUNISH_COST} token.
- Each deduction point you assign reduces THAT member's income this period by
  {PUNISH_IMPACT} tokens.
The other members decide at the same time and may assign deduction points to
anyone, including you. No one is told who assigned points to whom; you are only
told the total number of points you received.
{RECEIVED_LAST_PERIOD}
{SOCIETY_BLOCK}
{PERSONALITY_BLOCK}
DEDUCTIONS SO FAR
{PUNISH_HISTORY_BLOCK}

YOUR DECISION
Decide how many deduction points (0 to {MAX_PUNISH}) to assign to each member
shown above. You may reason briefly first. Then finish your reply with EXACTLY one
line in this format, and nothing after it:
DEDUCT: A=a B=b C=c
(where a, b, c are integers from 0 to {MAX_PUNISH}, for Member A, Member B,
Member C)
\end{Verbatim}
}

\clearpage
\printbibliography

\end{document}
